---
title: "Regression to the Mean"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading in libraries. I will use 'mvtnorm' to simulate two noramlly-distributed variables that are imperfectly correlated.
```{r, warning=FALSE}
library(mvtnorm)
library(scales)
library(tidyverse)
```


Let's make our two variables have a high correlation of 0.8 and then a low correlation of 0.2 and look at the difference. We will store these values in a vector that we will iterate across.
```{r}
covars <- c(0.8, 0.2)
```

For each of these covariances, or correlations between the two variables, we will simulate data and make a plot.
```{r, include=TRUE, warning=FALSE}
par(bty="n", mfrow=c(2,2)) # make a plot with no box type with 6 panels: 3 rows and 2 columns
for (covar in covars){
  
  covar_mat <- rbind(c(1,covar), c(covar,1))    # create covariance matrix

  # create table of draws from a bivariate normal distribution
  d <- rmvnorm(n=10000, mean = rep(0,2), sigma=covar_mat) %>%
    as_tibble() %>% rename("x" = V1, "y" = V2)
  
  # create another table that selects only extreme values for y 
  d_red <- d %>% 
    filter(y >= 2)
    
  # make a plot
  plot(d$x, d$y, 
     pch=16, 
     col=alpha("black", 0.1),
     main=paste("correlation = ", covar, sep=" "),
     xlab="x",
     ylab="y",
     axes=F, 
     ylim=c(-3.5, 3.5),
     xlim=c(-3.5, 3.5))
  points(d_red$x, 
    d_red$y, 
    col=alpha("red", 0.3), 
    pch=16)
  axis(side=1,  labels=T, at=c(-3,-2,-1,-1,0,1,2,3))
  axis(side=2, labels=T, at=c(-3,-2,-1,-1,0,1,2,3))
  abline(h=2, lty=3, col="blue", lwd=2)
  hist(d_red$x, col=alpha("red", 0.5), xlab="x", freq=F, main="", breaks=20)
  abline(v=2, lty=3, col="blue", lwd=2)
}
```

